Empirical normative theories

### 2020-08-04 simple.tex
Rm old abstract and intro

\begin{abstract}
	In decision theory and economy in general, preferences are generally considered as “intuitive”, as originating from an “immediate sensation”, as famously written by \citeauthor{von_neumann_theory_2004}. This conception is insufficient to study the positions that individuals adopt towards difficult decision problems after careful thinking. This article introduces, and defines formally, the concept of deliberated preferences to fill this gap: the deliberated preferences of an individual about a given decision problem represent the position of that individual towards that problem that is stable facing counter-arguments. As defined, deliberated preferences are only indirectly observable. To make it possible to model deliberated preferences, the article also defines formally the notion of a theory of deliberated preferences, together with a protocol of falsification of such a theory. The proposed falsification protocol involves only observations similar to those used in classical theories of revealed preferences. It is shown that a theory of deliberated preference captures effectively a known subset of the deliberated preferences if and only if it is not falsifiable.
\end{abstract}

\section{Motivation}
In decision theory and economy in general, preferences are generally considered as “intuitive”, as originating from an “immediate sensation”, following the terms used by \citet{von_neumann_theory_2004} (vNM) \citep{fishburn_retrospective_1989, cailloux_reasons_2017}. Though appropriate when studying the everyday behavior of individuals, this conception is insufficient to study the positions that individuals adopt towards difficult decision problems after careful thinking. In this article I define the concept of deliberated preferences to fill this gap: the deliberated preferences of an individual about a given decision problem represent the position of that individual towards that problem that is stable facing counter-arguments.

After having defined this concept formally, I define the notion of a theory pretending to capture the deliberated preferences of a set of individuals about some topic, I propose an empirical procedure to attempt to falsify such a theory, and I show that such a theory tells the truth whenever it is not falsifiable by that procedure.

This article does not attempt to propose a concrete theory of deliberated preferences. It is more preliminarly. It proposes formal definitions for 1) the kind of object that are to be studied for capturing “more reflexive” preferences than considered usually, 2) what theories of such preferences should look like, 3) how to falsify such theories and using which information. Therefore, it proposes a research program rather than a direct solution to the problem of capturing reflexive preferences.

Before delving into the formalism, let me present a bit of motivation for considering deliberated preferences as an interesting object to study.
Being able to capture deliberated preferences can help provide a better understanding of human preferences in the large sense and may improve policy making (though this article does not explore issues of group decision). Besides this very broad goal, here are three example of contexts in which this notion is useful, which may give it a more concrete interest and serve to introduce this concept informally by indicating the kinds of things that it should be able to help understand.

First, decision theory has produced several models of how an individual ought to decide, such as the seminal one of vNM. (The following remarks apply equally well to other normative models such as the one of \citet{savage_foundations_1972}, \citet{keeney_decisions_1993} or more recent ones.) Those models can be used to help individuals take “good” decisions, as opposed to predict their behavior (for which other models, taking into account cognitive limitations, are more appropriate \citep{wakker_prospect_2010}). But it is unclear how to validate recommendations obtained on the basis of such models. Some researchers have opposed the norms on which some of those models are based \citep{ellsberg_risk_1961, allais_so-called_1979}. As \citet{raiffa_back_1985} points out, while it is the case that some sophisticated decision makers may consciously refuse to abide by the vNM axioms (or others), it is also true that some decision makers simply fail to comply with them because of insufficient reflexion. However, no general proposal, as far as I know, has been described yet to test whether, for a given individual and context, those axioms would, after careful thinking, reveal acceptable. Psychologists have tried to test the convincing power of those axioms \citep{slovic_who_1974, maccrimmon_utility_1979}. Their conclusions inevitably depend on how axioms have been presented to the individuals and on whether counter-arguments have been presented, as \citet{slovic_who_1974} themselves point out. But lacking a systematic approach to determine stability facing counter-arguments, those conclusions only shed very partial light on the question of acceptance of specific norms by the general public. The notion of deliberated preferences can help confront different possible norms for normative decision theory by seeing them as constituents of theories of deliberated preferences and using the procedure described in this article to confront theories.

Second, the concept of nudge \citep{thaler_nudge:_2009} has been tremendously popular in the recent past, but has also been vastly criticized: the risk of manipulation is particularly high whenever it is unsure that the individual would indeed, after careful thinking, have adopted the decision that the nudge promotes.

Third, analytical philosophy has presented very elaborated and diverging views about the concept of justice. \citet{rawls_theory_1999}, in particular, suggests that the appropriate conception of justice is one that individuals would arrive at in a state of reflexive equilibrium (a concept borrowed from \citet{goodman_fact_1983}). \citeauthor{rawls_theory_1999} and \citeauthor{goodman_fact_1983} insist that individuals must not only accept principles, when thinking about difficult notions such as justice, but also think about consequences of those principles: sometimes, principles that appear intuitively acceptable yield consequences that do not appear intuitively acceptable, in which case something must give. The state of reflexive equilibrium is a state where accepted principles yield acceptable consequences only. The concept of deliberated preferences may be considered as defining that state of reflexive equilibrium, or to be a path towards such a goal (an important difference is that it does not explicitly model the difference between principles and consequences). One way to go beyond philosophers disagreement about the right conception of justice, then, is to ask non-philosophers about what they think is right and so use these laymen as arbiters of the competition between views of justice. Similar empirical studies have started to be conducted recently \citep{gaertner_empirical_2012}. But those studies are conducted by asking individuals only which distribution of goods they consider intuitively right in various circumstances. Thus, they do not try to capture their well-considered judgments for justice, where different reasonings in favor of different conceptions of justice would be presented to the individuals together with the distributions. Hence, this does not test appropriately whether laymen would accept, after thinking carefully, the sophisticated conceptions of justice presented by a given philosopher or economist. A more appropriate way of testing the convincing power of elaborated claims about justice may be considered to require, here again, to present his conception of justice to the individual and submit it to counter-arguments (possibly coming from other sophisticated thinkers).

The first and third contexts of usage come together when it is desired to test whether an individual accepts some axiom as representing an acceptable way of reasoning for him. One simple way is to explain axioms individually and ask the individual whether they appear acceptable. But this fails to make sure the individual understands and accepts the consequences of the union of the axioms accepted by such a procedure \citep{meinard_justification_2018}. (Note that probably nobody is able to mentally deduce all consequences of a union of non trivial axioms.) As an example, in a context of trying to determine the voting rule that best represents someone’s notion of justice, the tested individual could accept the axioms of universal domain, independence of irrelevant alternatives, and pareto dominance, without realizing that he has just accepted dictatorship \citep{arrow_social_1963}.

>>>>

\begin{abstract}
	TODO
\end{abstract}

\section{Motivation}
